---
title: "Using Shiny App Builder"
author: "Jenna Reps, Jamie Gilbert, Josh Ide"
date: "`r Sys.Date() `"
output: html_document
vignette: >
  %\VignetteIndexEntry{guide}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using the Shiny App Builder

The shiny app builder provides a way to combine different shiny modules into a single app.  For example, if you have a characterization study, an estimation study and a prediction study that are related as they all use the same cohorts, then you may want to view the results in a single shiny app.  This can be done by using `ShinyAppBuilder` to combine the characterization, estimation and prediction shiny modules in `OhdsiShinyModule`. The main source of shiny modules is the `OhdsiShinyModule` R package, however, it is possible to add modules from other R packages.

In this vignette we provide examples on how to use the `ShinyAppBuilder` to create flexible shiny apps for exploring OHDSI results.

The `ShinyAppBuilder` requires that all results to be explored by the shiny app are saved into a single database (i.e., all results for the different shiny modules in an app are saved into the same database), as a single database connection is shared across shiny modules. 

# Example 1

To create a shiny app that contains four shiny modules:

- About (an introduction to the shiny app)
- Cohort Generation (details about generated cohorts)
- Cohort Method (details about causal inference studies via the cohort method)
- Prediction (details about patient level prediction models)

All of these are available as shiny modules in `OhdsiShinyModules`.  

## Creating module config settings 
To create the shiny app via `ShinyAppBuilder` we first need to create a config specification for all the shiny modules we wish to include into the single shiny app.  A config can be created using `createModuleConfig`.  

```{r echo=FALSE}

inputsForConfig <- data.frame(
  Inputs = c('moduleId','tabName','shinyModulePackage',
      'moduleUiFunction', 'moduleServerFunction',
      'moduleDatabaseConnectionKeyService',
      'moduleDatabaseConnectionKeyUsername',
      'moduleInfoBoxFile',
      'moduleIcon','resultDatabaseDetails','useKeyring'
),

Description = c("a unique id for the shiny app", "The menu text for the module", "The R package that contains the shiny module", "The name of the module's UI function", "The name of the module's server function", 
                "A name for the keyring service/environment variable to store the result database details",
                "A name for the keyring username/environment variable to store the result database details",
                "The function in the shinyModulePackage package that contains the helper information",
                "An icon to use in the menu for this module",
                "A list containing the results details: dbms, tablePrefix, cohortTablePrefix, databaseTablePrefix, schema, databaseTable, incidenceTablePrefix",
                "Whether to use keyring of environmental variables")
)

knitr::kable(inputsForConfig)

```

Note: it is possible to add shiny modules from any R package by setting `shinyModulePackage` to the R package with the UI and server functions and then specifying the UI function as `moduleUiFunction` and server function as `moduleServerFunction`.  However, the server function must take as input `id` (the module id as standard for shiny server modules) and `resultDatabaseSettings` (a list containing the database result details required when extracting the results from the database).

## Creating the about module config 

For the about module we will use the about shiny modules in `OhdsiShinyModule`.  The UI is named `aboutViewer`, the server is named `aboutServer` and the about helper function is called `aboutHelperFile()`. The about module does not have any results, so `moduleDatabaseConnectionKeyService`, `moduleDatabaseConnectionKeyUsername` and `resultDatabaseDetails` are NULL.  As the about module provides information about the shiny app, the 'info' icon seems appropriate. The inputs into `createModuleConfig` for an about module are:

```{r eval=FALSE}

aboutModule <- createModuleConfig(
      moduleId = 'about',
      tabName = "About",
      shinyModulePackage = "OhdsiShinyModule",
      moduleUiFunction = 'aboutViewer',
      moduleServerFunction = 'aboutServer',
      moduleDatabaseConnectionKeyService = NULL,
      moduleDatabaseConnectionKeyUsername = NULL,
      moduleInfoBoxFile =  "aboutHelperFile()",
      moduleIcon = 'info',
      resultDatabaseDetails = NULL,
      useKeyring = TRUE
    )

```

For simplicity, the `ShinyAppBuilder` contains a function called `createDefaultAboutConfig` with these default about settings into `createModuleConfig`, this is quicker to use than `createModuleConfig` if you are using a standard about module.  Alternatively, you could just run:

```{r eval=FALSE}

aboutModule <- createDefaultAboutConfig(
      resultDatabaseDetails = NULL,
      useKeyring = TRUE
    )

```

## Creating the prediction module config 

To add a prediction module you can the `OhdsiShinyModule` functions: `aboutPrediction` for the module UI, `aboutPrediction` for the module server and `aboutPredictionFile()` for the about helper function. A suitable icon is `chart-line`.  For the prediction module, results in the database format created by the `PatientLevelPrediction` package must be in a database that will be connected to when viewing the shiny app.  The `resultDatabaseDetails` input is a list that must specify the details about the result tables:

- dbms (the database management system with the results e.g., 'sqlite')
- tablePrefix (a string that is prefixed to the prediction result tables e.g., 'plp_')
- cohortTablePrefix (a string that is prefixed to the cohort table result e.g., 'cg_')
- databaseTablePrefix (a string that is prefixed to the database table result e.g., '')
- schema (the schema with the result tables e.g., 'main')
- databaseTable (the name of the database table e.g., 'DATABASE_META_DATA')

```{r eval=FALSE}

resultDatabaseDetails <- list(
      dbms = 'sqlite',
      tablePrefix = 'plp_',
      cohortTablePrefix = 'cg_',
      databaseTablePrefix = '',
      schema = 'main',
      databaseTable = 'DATABASE_META_DATA'
    )

predictionModule <- createModuleConfig(
    moduleId = 'prediction',
    tabName = "Prediction",
    shinyModulePackage = 'OhdsiShinyModules',
    moduleUiFunction = "predictionViewer",
    moduleServerFunction = "predictionServer",
    moduleDatabaseConnectionKeyService ="resultDatabaseDetails",
    moduleDatabaseConnectionKeyUsername = "prediction",
    moduleInfoBoxFile =  "predictionHelperFile()",
    moduleIcon = "chart-line",
    resultDatabaseDetails = resultDatabaseDetails,
    useKeyring = TRUE
    )

```

For simplicity, the `ShinyAppBuilder` contains a function called `createDefaultPredictionConfig` with these default prediction settings, however, you must still specify whether you want to use keyring `useKeyring` and you must specify the `resultDatabaseDetails`.  Atlernatively, you could just run:

```{r eval=FALSE}

resultDatabaseDetails <- list(
      dbms = 'sqlite',
      tablePrefix = 'plp_',
      cohortTablePrefix = 'cg_',
      databaseTablePrefix = '',
      schema = 'main',
      databaseTable = 'DATABASE_META_DATA'
    )

predictionModule <- createDefaultPredictionConfig(
      resultDatabaseDetails = resultDatabaseDetails,
      useKeyring = TRUE
    )

```

## Creating the cohort generation and cohort method using default functions

We have default config creation for cohort method and cohort generation in `ShinyAppBuilder`:

```{r eval=FALSE}

resultDatabaseDetails <- list(
      dbms = 'sqlite',
      tablePrefix = 'cm_',
      cohortTablePrefix = 'cg_',
      databaseTablePrefix = '',
      schema = 'main',
      databaseTable = 'DATABASE_META_DATA'
    )

cohortMethodModule <- createDefaultEstimationConfig(
      resultDatabaseDetails = resultDatabaseDetails,
      useKeyring = TRUE
    )

cohortGeneratorModule <- createDefaultCohortGeneratorConfig()

```

## Combining config settings

Next step is to combine the module config settings into a shiny app config.  First we use `initializeModuleConfig()` to create an empty shiny app config and then we use `addModuleConfig()` to add each of the module configs we previously created:


```{r eval=FALSE}

library(dplyr)
shinyAppConfig <- initializeModuleConfig() %>%
  addModuleConfig(aboutModule) %>%
  addModuleConfig(cohortGeneratorModule) %>%
  addModuleConfig(cohortMethodModule) %>%
  addModuleConfig(predictionModule)

```

It is possible to save the shiny app config using `saveConfig(shinyAppConfig, 'save location')` and load a previously saved shiny app config `shinyAppConfig <- loadConfig('save location')`

## View Shiny App

### Create a shiny app for a shiny server

To run the shiny app on a shiny server, specify the connection details to the result database and then just add the following lines of code

```{r eval=FALSE}
# create a connection to the result database
connectionDetails <- DatabaseConnector::createConnectionDetails()
createShinyApp(shinyAppConfig, connectionDetails = connectionDetails)
```


### View shiny app

To just view the shiny app locally, specify the connection details to the result database and then just add the following lines of code

```{r eval=FALSE}
# create a connection to the result database
connectionDetails <- DatabaseConnector::createConnectionDetails()
viewShiny(shinyAppConfig, connectionDetails = connectionDetails)
```

# Example 1 Full Code To View Locally

```{r eval=FALSE}
library(ShinyAppBuilder)
library(dplyr)

aboutModule <- createModuleConfig(
      moduleId = 'about',
      tabName = "About",
      shinyModulePackage = "OhdsiShinyModule",
      moduleUiFunction = 'aboutViewer',
      moduleServerFunction = 'aboutServer',
      moduleDatabaseConnectionKeyService = NULL,
      moduleDatabaseConnectionKeyUsername = NULL,
      moduleInfoBoxFile =  "aboutHelperFile()",
      moduleIcon = 'info',
      resultDatabaseDetails = NULL,
      useKeyring = TRUE
    )

resultDatabaseDetails <- list(
      dbms = 'sqlite',
      tablePrefix = 'plp_',
      cohortTablePrefix = 'cg_',
      databaseTablePrefix = '',
      schema = 'main',
      databaseTable = 'DATABASE_META_DATA'
    )

predictionModule <- createModuleConfig(
    moduleId = 'prediction',
    tabName = "Prediction",
    shinyModulePackage = 'OhdsiShinyModules',
    moduleUiFunction = "predictionViewer",
    moduleServerFunction = "predictionServer",
    moduleDatabaseConnectionKeyService ="resultDatabaseDetails",
    moduleDatabaseConnectionKeyUsername = "prediction",
    moduleInfoBoxFile =  "predictionHelperFile()",
    moduleIcon = "chart-line",
    resultDatabaseDetails = resultDatabaseDetails,
    useKeyring = TRUE
    )

resultDatabaseDetails <- list(
      dbms = 'sqlite',
      tablePrefix = 'cm_',
      cohortTablePrefix = 'cg_',
      databaseTablePrefix = '',
      schema = 'main',
      databaseTable = 'DATABASE_META_DATA'
    )

cohortMethodModule <- createDefaultEstimationConfig(
      resultDatabaseDetails = resultDatabaseDetails,
      useKeyring = TRUE
    )

cohortGeneratorModule <- createDefaultCohortGeneratorConfig()


# add the modules into a single shiny config
shinyAppConfig <- initializeModuleConfig() %>%
  addModuleConfig(aboutModule) %>%
  addModuleConfig(cohortGeneratorModule) %>%
  addModuleConfig(cohortMethodModule) %>%
  addModuleConfig(predictionModule)


# add connection details to result database
connectionDetails <- DatabaseConnector::createConnectionDetails()
viewShiny(shinyAppConfig, connectionDetails = connectionDetails)

```
